---
title: "Explore MB's Economic Data"
output: html_document
---

# Setup

Notes:

* `CTRL + ALT + i` to insert R "chunk".  
* Name R chunks to see them in "Document outline" (right)
* To run chunk, hit green arrow or (keyboard shortcut) `CMD + SHIFT +ENTER`

```{r libraries}
# Reading Data
library(readr)

# Processing Data
library(dplyr)
library(tibble)
library(tidyr)
library(lubridate)

# Visualizing Data
library(ggplot2) # Loads "diamonds" dataset"
```

```{r directories}
project_directory <- "C:/Users/Ryan/repos/792data"
```


# Load Data

Notes:

* You can also use the `xlsx` package to load in individual sheets from an Excel file. You will need to install `rJava` as well as Java.
* `read_csv` is from `readr`. Better defaults. Converts characters to characters not "factors". `read.csv` (dot) will convert variables to factors by default when reading in a CSV file. 
* Factors are technically integers but display (print) as characters.

```{r, message=FALSE}
hours_data <- read_csv(file = paste0(project_directory,
                                     "/data/raw/or_2017_dataset_hours_sheet.csv"))
```

# Process Data

What are we seeing? The first column "X1" is obviously a date variable. Make this clear and convert to a date.

```{r}
glimpse(hours_data)
```

Note:

* Example of how to join data
* To comment out, `CTRL + SHIFT + C`

```{r}
# combined_data <- left_join(x = hours_data,
#           y = other_data,
#           by = c("left_var" = "right_var", 
#                  "left2" = "right2"))
```

```{r}
hours_data <- hours_data %>% 
  # Fix variable name (excel default)
  rename(Hours_year_str = X1) %>% # Hours_year => new name. X1 => old name.
  # Create month/day/year variables, keep old
  separate(col = Hours_year_str, 
           into = c("Hour_month", "Hour_day", "Hour_year"),
           sep = "[/]",
           remove = FALSE) %>%
  # Create new date variable manually
  mutate(Hours_year_manual = mdy(paste0(Hour_month,
                                       "01",
                                       Hour_year,
                                       sep = "[/]"))) %>% 
  # Fix issue where 1960 becomes 2060 (other years too)
  mutate(Hours_year_manual = if_else(year(Hours_year_manual) >= 2060, 
                              Hours_year_manual - years(100),
                              Hours_year_manual))
```

# Analysis

